---
title: "Mandatory Assignments in Econometrics"
author: "Hege Johansen"
format: html
editor: visual
---

# Mandatory Assignment 1

```{r setup}
#| echo: false
#| output: false

library(tidyverse)
library(tinytex)
library(dplyr)
library(ggplot2)
library(dineq)
library(readxl)
library(magrittr)
library(vtable)
```

```{r EU-data}
#| echo: false
#| output: false

# EU data taken from: 
  # https://ec.europa.eu/eurostat/databrowser/view/demo_r_pjanaggr3/default/table?lang=en
  # https://ec.europa.eu/eurostat/databrowser/view/nama_10r_3gdp/default/table?lang=en
 

eu_pop_nuts3 <- read_excel("demo_r_pjanaggr3__custom_3572346_spreadsheet.xlsx", 
    sheet = "cleaned_pop")
eu_gdp_nuts3 <- read_excel("nama_10r_3gdp__custom_3572365_spreadsheet.xlsx", 
    sheet = "cleaned_gdp")
```

```{r UK-data}
#| echo: false
#| output: false

# Uk data taken from:
  # https://www.ons.gov.uk/economy/grossdomesticproductgdp/datasets/regionalgrossdomesticproductallnutslevelregions

  uk_pop <- read_excel("regionalgrossdomesticproductgdpallitlregions.xlsx", 
      sheet = "Table 6", skip = 1)
  
  uk_gdp <- read_excel("regionalgrossdomesticproductgdpallitlregions.xlsx", 
      sheet = "Table 5", skip = 1)
   
  # Drop non nuts 3 regions
   uk_gdp_nuts3 <- uk_gdp %>%
    # Count the numbers of id characters 
    mutate("N_car_id" = nchar(uk_gdp$`ITL code`)) %>%
    # Keep thoes that have 5 e.g. NUTS3 Regions
   filter(N_car_id == 5)
   
   uk_pop_nuts3 <- uk_pop %>%
    # Count the numbers of id characters 
    mutate("N_car_id" = nchar(uk_pop$`ITL code`)) %>%
    # Keep thoes that have 5 e.g. NUTS3 Regions
    filter(N_car_id == 5) 

```

```{r EU-popdata}
#| echo: false
#| output: false

# Transformation from wide to long
 eu_pop_nuts3 <- eu_pop_nuts3 %>% 
   mutate_if(is.numeric,as.character, is.factor, as.character) %>% 
   pivot_longer(
    cols = c('2000':'2020'),
    names_to = "Year",
    values_to = "Population"
  )
 
  # Renaming geo column
      eu_pop_nuts3 <- rename(eu_pop_nuts3, "Region_code" = "GEO (Codes)")
```

```{r UK-pop-data}
#| echo: false
#| output: false

# Transformation from wide to long
  uk_pop_nuts3 <- uk_pop_nuts3 %>% 
   mutate_if(is.numeric,as.character, is.factor, as.character) %>% 
   pivot_longer(
    cols = c('2000':'2020'),
    names_to = "Year",
    values_to = "Population"
    )
  
# Renaming a column
   uk_pop_nuts3 <- rename(uk_pop_nuts3, "Region_code" = "ITL code")
# Keep relevant vars
   uk_pop_nuts3 <- uk_pop_nuts3 %>% select("Region_code", "Year","Population")    

```

```{r Merge of UK and EU data}
#| echo: false
#| output: false

pop_all_nuts3 <- rbind(uk_pop_nuts3, eu_pop_nuts3)

```

```{r EU-gdp-data}
#| echo: false
#| output: false

# Transforming from wide to short
  eu_gdp_nuts3 <- eu_gdp_nuts3 %>%  
  mutate_if(is.numeric,as.character, is.factor, as.character) %>%  
  pivot_longer(
    cols = c('2000':'2020'),
    names_to = "Year",
    values_to = "GDP")
  
# Renaming geo column
      eu_gdp_nuts3 <- rename(eu_gdp_nuts3, "Region_code" = "GEO (Codes)")

```

```{r UK-gdp-data}
#| echo: false
#| output: false

# Transforming from wide to short
  uk_gdp_nuts3 <- uk_gdp_nuts3 %>% 
  mutate_if(is.numeric,as.character, is.factor, as.character) %>%  
  pivot_longer(
    cols = c('2000':'2020'),
    names_to = "Year",
    values_to = "GDP")
  
# Transforming from GDP in pounds to GDP in Euro
uk_gdp_nuts3 <- uk_gdp_nuts3 %>% 
  mutate("GDP" = c(as.numeric(GDP) * 1.14))

# Renaming geo column
    uk_gdp_nuts3 <- rename(uk_gdp_nuts3, "Region_code" = "ITL code")
      
# Keep relevant vars
    uk_gdp_nuts3 <- uk_gdp_nuts3 %>% select("Region_code", "Year", "GDP")

```

```{r Merge of UK- and EU-gdp-data}
#| echo: false
#| output: false   

gdp_all_nuts3 <- rbind(uk_gdp_nuts3, eu_gdp_nuts3 )

```

```{r Conversion from GDP in millions to GDP in ones}
#| echo: false
#| output: false

gdp_all_nuts3 <- gdp_all_nuts3 %>% 
  mutate("GDP" = as.numeric(GDP) * 1000000) 

```

```{r Merging of gdp- and pop-tables}
#| echo: false
#| output: false

GDPandPOP_all_nuts3 <- merge(gdp_all_nuts3, pop_all_nuts3)

```

```{r Removal of rows with missing values}
#| echo: false
#| output: false

GDPandPOP_all_nuts3 <- na_if(GDPandPOP_all_nuts3, ":") %>% 
   na.omit(GDPandPOP_all)

```

## Calculation of GDP per Capita

The formula for calculating GDP per Capita is as follows:

$y_i=GDP_i/population_i$

```{r Calculation GDP per Capita-column}

# Calculation of GDP per Capita for all regions
GDPperCap_all_nuts3 <- mutate(GDPandPOP_all_nuts3, "GDP_per_capita" = as.numeric(GDPandPOP_all_nuts3$GDP) / as.numeric(GDPandPOP_all_nuts3$Population))
```

```{r column for Nuts 1 id}
#Gen Nuts 1 id
GDPperCap_all_nuts3 %<>%  
  mutate(id_nuts0 = substr(GDPperCap_all_nuts3$"Region_code", 1,2))

#Gen Nuts 3 id
GDPperCap_all_nuts3 %<>%  
  mutate(id_nuts3 = substr(GDPperCap_all_nuts3$"Region_code", 1, 5))
  
```

### GDP per Capita by country

```{r dataframe for Finland and Turkey}
#| echo: false
#| output: false

Finland_nuts3 <- GDPperCap_all_nuts3 %>% 
  filter(grepl("FI", id_nuts0))
Turkey_nuts3 <- GDPperCap_all_nuts3 %>% 
  filter(grepl("TR", id_nuts0))

FI_TR_nuts3 <- rbind(Finland_nuts3, Turkey_nuts3)
```

```{r}
#| echo: false

FI_TR_nuts3 %>%  
  distinct(id_nuts0, Year, .keep_all = TRUE) %>%  

  ggplot(aes(x = Year,
        y = GDP_per_capita, 
        fill = id_nuts0, 
        color = id_nuts0,
        group = id_nuts0)) +
  geom_line(lwd = 0.8) +
  labs(x = "Year", y = "GDP per capita") +
  ggtitle("Changes in GDP per Capita by country") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

### GDP per Capita - Finland

```{r GDP per capita for Finland regions}
#| echo: false

Finland_nuts3 %>%  
  distinct(Region_code, Year, .keep_all = TRUE) %>% 
  ggplot(aes(x = Year, 
        y = GDP_per_capita, 
        fill = Region_code, 
        color = Region_code,
        group = Region_code)) +
  geom_line(lwd = 0.8) +
   labs(x = "Year", y = "GDP per capita") +
  ggtitle("Finland: GDP per capita - NUTS3") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

#### The 3 wealthiest regions 2019

```{r wealthiest regions in finland 2019}
#| echo: false

FI_TR_nuts3 %>%
  filter(grepl("FI...", id_nuts3), Year == 2019) %>%
  select(id_nuts3, GDP_per_capita) %>%
  slice_max(GDP_per_capita, n = 3)
```

FI1B1 = Helsinki, FI200 = Åland, FI 195 = Pohjanmaa

#### The 3 least wealthiest regions in 2019

```{r least wealthiest regions in finland 2019}
#| echo: false

FI_TR_nuts3 %>%
  filter(grepl("FI...", id_nuts3), Year == 2019) %>%
  select(id_nuts3, GDP_per_capita) %>%
  slice_min(GDP_per_capita, n = 3)

```

FI1D1 = Etelä-Savo, FI1C3 = Päijät-Häme, FI1D8 = Kainuu

### GDP per Capita - Turkey

```{r GDP per capita for Turkey regions}
#| echo: false

Turkey_nuts3 %>%  
  distinct(Region_code, Year, .keep_all = TRUE) %>% 
  ggplot(aes(x = Year, 
        y = GDP_per_capita, 
        fill = Region_code, 
        color = Region_code,
        group = Region_code)) +
  geom_line(lwd = 0.8) +
   labs(x = "Year", y = "GDP per capita") +
  ggtitle("Turkey: GDP per capita - NUTS3") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

#### The 3 wealthiest regions 2019

```{r wealthiest regions in Turkey 2019}
#| echo: false

FI_TR_nuts3 %>%
  filter(grepl("TR...", id_nuts3), Year == 2019) %>%
  select(id_nuts3, GDP_per_capita) %>%
  slice_max(GDP_per_capita, n = 3)
```

TR100 = Instanbul, TR421 = Kocaeli, TR510 = Ankara

#### The 3 least wealthiest regions in 2019

```{r Turkeys least wealthiest regions 2019}
#| echo: false

FI_TR_nuts3 %>%
  filter(grepl("TR...", id_nuts3), Year == 2019) %>%
  select(id_nuts3, GDP_per_capita) %>%
  slice_min(GDP_per_capita, n = 3)
```

TRA21 = Agri, TRC21 = Sanliurfa, TRB21 = Van

## Regional inequity

### Calculation of the population watertight GDP Ginie coefficients for the European NUTS2 () level.

$$GINW_j=\frac{1}{2 \bar{y_j}} \sum_{i}^{n_j}\sum_{l}^{n_j}\frac{p_i}{P_j} \frac{p_l}{P_j} |y_i-y_l|$$

```{r}
GDPperCap_all_nuts3$GDP_per_capita <- as.numeric(GDPperCap_all_nuts3$GDP_per_capita)
GDPperCap_all_nuts3$Population <- as.numeric(GDPperCap_all_nuts3$Population)

gini.wtd(GDPperCap_all_nuts3$GDP_per_capita, weights = GDPperCap_all_nuts3$Population)
```

```{r}
#Nuts 2 id
GDPperCap_all_nuts3 <- GDPperCap_all_nuts3 %>%  
  mutate(id_nuts2 = substr(GDPperCap_all_nuts3$`Region_code`,1,4))
#Nuts 2 id
GDPperCap_all_nuts3 <- GDPperCap_all_nuts3 %>%  
  mutate(id_nuts0 = substr(GDPperCap_all_nuts3$`Region_code`,1,2))
```

```{r}
GDPperCap_all_nuts3  %<>% 
  group_by(id_nuts2) %>% 
  mutate(gini_n2 = gini.wtd(GDP_per_capita, w = Population)) %>%
  ungroup() # to remove grouping
```

## Gini-coefficients for NUTS2 regions - Czech Republic

```{r}

```

## Gini-coefficients for NUTS2 regions - Finland

## Gini-coefficients for NUTS2 regions - United Kingdom

## Gini-coefficients for NUTS2 regions - Turkey

```{r GDP per capita for Turkey regions}
#| echo: false

 Turkey_gini_nuts2 <- GDPperCap_all_nuts3 %>%  
  distinct(Region_code, Year, .keep_all = TRUE) 
```

```{r}
   gini_turkey <- Turkey_gini_nuts2 %>%  
     ggplot(aes(x = Year, 
        y = gini_n2, 
        fill = Region_code, 
        color = Region_code,
        group = Region_code)) +
  geom_line(lwd = 0.8) +
   labs(x = "Year", y = "GDP per capita") +
  ggtitle("Turkey: GDP per capita - NUTS3") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
   View(gini_turkey)
```

### Data description

### The distribution of the Ginie coefficients

```{r}
#| echo: false
#| output: true
hist_gini_n2 <- ggplot(
  GDPperCap_all_nuts3,
  mapping = aes(x = gini_n2)) +
  geom_histogram(
    colour = "white", 
    fill = "green", 
    bins = 20, 
    alpha = 0.5)

 density_gini_n2 <- ggplot(
    data = GDPperCap_all_nuts3,
    mapping = aes(x = gini_n2 )) +
  geom_density(
    alpha = 0.2, 
    fill = "green")

gridExtra::grid.arrange(
  hist_gini_n2, 
  density_gini_n2,
  nrow = 2)
```

# Mandatory Assignment 2
