---
title: "Mandatory assignment 2"
author: "Hege Johansen"
format: html
editor: visual
bibliography: [Econometrics-Hege.bib]
csl: apa-no-ampersand.csl.
---

## Growth and inequity

### Data

### **Estimate the effete of regional development** $y_j$ **on regional inequality** $GINIW_j$ **for the year 2010**

Economical development is the logarithm of GDP per capita, according to [@lessmann2017].
In this case that is regional development.

```{r echo=FALSE}

# creating new dataframe for 2010 - nuts 2
data2010_nuts2 <- GDPperCap_all_nuts3 %>% 
  filter(Year == 2010) %>% 
  select(GDP_per_capita, id_nuts2, gini_n2) %>% 
  mutate(log_gdp_pc = log(GDP_per_capita))
```

```{r}


# estimate for all countries
lm_allcountries <- lm(gini_n2~log_gdp_pc, data2010_nuts2) 

lm_allcountries
```

### **Discuss the goodness of fit of your estimated model.**

### **Plot the relationship between regional development** $y_j$ **and regional inequality** $GINIW_j$ **and the fitted line corresponding to your estimate.**

```{r}
# Save coefficients
b1_allcountries <- lm_allcountries$coefficients[1]
b2_allcountries <- lm_allcountries$coefficients[2]

data2010_nuts2 %>% 
  ggplot() +
  xlim(0, 15) +
  ylim(-0.3, 0.5) +
  geom_point(aes(x = log_gdp_pc, y = gini_n2)) +
  geom_abline(intercept = b1_allcountries, slope = b2_allcountries, colour = "green", size = 1) +
 theme_minimal()
```

### **Plot the residuals against the predicted values of your model.**

### **Discuss the classical assumptions OLS in light of your data and plots**

## Other determinants of inequity

### Data

Download from [Eurostat](https://ec.europa.eu/eurostat/web/regions/data/database) for you subset of countries regional (NUTS2, j ) data related to transport infrastructure, education and demographics.
Select on variable per category that you would like to explore further in there relationship to regional inequality $GINIW_j$ .

### Estimate a multiple linear regression model with you new variables for 2010 and give a small interpretation of your findings.

### Discuss the overall fit of your model

### Discuss the inference related to your findings
